---
title: "First look at confluence behavior, Loire headwaters"
author: "Jacob Diamond, Gilles Pinay, Florentina Moatar et al."
date: "24/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, cache=FALSE)
```



Simple intro analysis: are downstream DO values within the range of upstream values
on any given day? We focus here on the on daily minima and maxima as they reflect GPP 
and ER, respectively. For example, if the downstream 
minimum DO on day 1 was 8 mg/L, and the minima for the two upstream sites were 
8.5 mg/L and 9 mg/L, then the downstream minimum would be "below" the range constrained
by the upstream sites. This would imply non-conservative mixing behavior.

The sites below are arranged from left-to-right by decreasing difference in 
confluence tributary specific conductance. i.e., the Fontbonne-Violay confluence
has the greatest relative specific conductivity difference (145%) and the 
Coise-Coizet has the smallest relative difference (1%).

```{r, echo=FALSE}
# Set working directory
setwd("Z:/Loire_DO")

# Load libraries
library(lubridate)
library(readxl)
library(transformr)
library(sf)
library(gganimate)
library(ggthemes)
library(ggspatial)
library(plotly)
library(tidyverse)

# Load data
df_DO <- readRDS("Headwaters/Data/DO_time_series_wide") %>%
  mutate(DOsat = if_else(DO_temp == 0,
                        0,
                        14.652 - 0.41022 * DO_temp + 0.007991 * 
                          DO_temp^2 - 0.000077774 * DO_temp^3),
         DO_per = DO * 100/ DOsat)

# Get rid of bad data when sensors were buried or when out of water
flood_sites1 <- c("Loise Feurs", "Mare Pont du diable", "Mare aval",
                 "Vizézy amont Bullieux", "Toranche aval", "Mare Azieux")
flood_sites2 <- c("Loise aval Poncins", "Loise Essertine en Donzy")
df_DO <- df_DO %>%
  filter(!(Site %in% flood_sites1 & between(datetime, 
                                            ymd_hm("2019-08-06 16:00"),
                                            ymd_hm("2019-08-29 16:45"))),
         !(Site %in% flood_sites2 & between(datetime, 
                                            ymd_hm("2019-08-06 16:00"),
                                            ymd_hm("2019-08-30 10:30"))),
         !(Site == "Coise aval Montrond" & between(datetime, 
                                                   ymd_hm("2019-08-06 16:00"),
                                                   ymd_hm("2019-09-11 11:30"))),
         !(Site == "Loise amont Doise Salt" & between(datetime, 
                                                   ymd_hm("2019-07-16 00:30"),
                                                   ymd_hm("2019-07-20 20:00"))),
         !(Site == "Loise aval Doise Salt" & between(datetime, 
                                                      ymd_hm("2019-08-06 19:00"),
                                                      ymd_hm("2019-08-30 10:15"))),
         !(Site == "Loise Essertine en Donzy" & between(datetime, 
                                                   ymd_hm("2019-07-20 22:15"),
                                                   ymd_hm("2019-07-23 15:30"))),
         !(Site == "Vizézy amont Bullieux" & between(datetime, 
                                                   ymd_hm("2019-07-08 19:00"),
                                                   ymd_hm("2019-07-16 14:15"))),
         !(Site == "Vizézy amont Bullieux" & between(datetime, 
                                                   ymd_hm("2019-07-24 23:00"),
                                                   ymd_hm("2019-08-29 15:30"))),
         !(Site == "Toranche Pontet" & between(datetime, 
                                               ymd_hm("2019-07-12 19:45"),
                                               ymd_hm("2019-07-21 17:30"))),
         !(Site == "Toranche Pontet" & between(datetime, 
                                               ymd_hm("2019-07-23 22:45"),
                                               ymd_hm("2019-07-29 01:00"))),
         !(Site == "Toranche Pontet" & between(datetime, 
                                               ymd_hm("2019-07-31 03:45"),
                                               ymd_hm("2019-08-06 17:30"))),
         !(Site == "Doise" & between(datetime, 
                                     ymd_hm("2019-07-14 15:00"),
                                     ymd_hm("2019-07-20 17:15"))),
         !(Site == "Doise" & between(datetime, 
                                     ymd_hm("2019-08-03 20:15"),
                                     ymd_hm("2019-08-06 12:30")))
         
         )

# Use threshold of daily temperature and %sat
df_DO_sum <- df_DO %>%
  mutate(date = date(datetime)) %>%
  group_by(Site,date) %>%
  summarize(do_mean = mean(DO_per, na.rm = TRUE),
            do_amp = max(DO_per, na.rm = T) - min(DO_per, na.rm = T),
            wat_t = max(DO_temp, na.rm = T) - min(DO_temp, na.rm = T)) %>%
  mutate(oow = if_else((do_mean > 90 & wat_t > 5) |
                         (do_mean > 90 & do_amp < 2) |
                         (wat_t > 8), "yes", "no"),
         buried = if_else(do_mean < 10, "yes", "no"))

# filter days that were out of water
df_DO <- df_DO %>%
  mutate(date = date(datetime)) %>%
  left_join(df_DO_sum) %>%
  filter(oow == "no", buried == "no")

# add a column for west/east
side <- tibble(Watershed = c("Coise", "Loise", "Toranche", "Lignon", "Mare"),
               side = c("east", "east", "east", "west", "west"))

# Read in watershed info
ws_meta <- read_excel("Headwaters/Data/Field measurements/sensor_metadata.xlsx",
                      sheet = 7)

df_DO <- select(ws_meta, Site = site_sensor, position, conf) %>%
  right_join(df_DO) %>%
  rename(watershed = Watershed, subwatershed = Subwatershed, area = area_km2, site= Site) %>%
  mutate(side = fct_rev(side)) %>%
  arrange(watershed, area) %>%
  mutate(site_code = fct_inorder(site_code),
         site = fct_inorder(site))

# Find days when the downstream site daily max or min is not within the 
# bounds of the upstream sites
df_daily <- df_DO %>%
  filter(!is.na(position)) %>%
  mutate(date = date(datetime)) %>%
  group_by(site, date, conf, position) %>%
  summarize(do_min = min(DO, na.rm = TRUE),
            do_max = max(DO, na.rm = T))

# Calculate the theoretical mixing based on conductivity 
df_cond <- df_DO %>%
  filter(!is.na(cond)) %>%
  select(site, date, conf, position, datetime, cond) %>%
  group_by(site, date, conf, position) %>%
  summarize(cond = mean(cond, na.rm = TRUE)) %>%
  ungroup() %>%
  select(-site) %>%
  group_by(conf, date) %>%
  pivot_wider(names_from = position, values_from = cond) %>%
  filter(map(up, length) > 1) %>% #filter for days where both upstreams have a value 
  drop_na() %>%
  mutate(ratio = map2(up, down, ~(.y - pluck(.x, 1)) / (pluck(.x, 2) - .y)))

# now test
df_test <- df_daily %>%
  ungroup() %>%
  select(-site) %>%
  pivot_wider(names_from = position, values_from = c(do_min, do_max)) %>%
  filter(map(do_min_up, length) > 1) %>% #filter for days where both upstreams have a value 
  drop_na() %>%
  mutate(min_out = map2(do_min_down, do_min_up, ~findInterval(.x, sort(.y))),
         max_out = map2(do_max_down, do_max_up, ~findInterval(.x, sort(.y)))) %>%
  mutate(min_out = if_else(min_out == 1, "in",
                           if_else(min_out == 0, "below", "above")),
         max_out = if_else(max_out == 1, "in",
                           if_else(max_out == 0, "below", "above"))) %>%
  mutate(per_ab = if_else(max_out == "above",
                          map2(do_max_down, do_max_up, ~(.x - max(.y)) / max(.y)),
                          list(NA_real_)),
         per_be = if_else(min_out == "below",
                          map2(do_min_down, do_min_up, ~(.x - min(.y)) / min(.y)),
                          list(NA_real_)))

# Combine estimated ratios with DO
df_test2 <- left_join(df_cond, df_test) %>%
  mutate(ratio_min = map2(do_min_up, do_min_down, ~(.y - pluck(.x, 1)) / (pluck(.x, 2) - .y)),
         ratio_max = map2(do_max_up, do_max_down, ~(.y - pluck(.x, 1)) / (pluck(.x, 2) - .y))) %>%
  select(date, conf, ratio, ratio_min, ratio_max)

# Plot it
df_test %>%
  select(date, conf, min_out, max_out) %>%
  pivot_longer(-c(conf,date)) %>%
  mutate(name = recode(name, 
                       max_out = "downstream daily maximum DO",
                       min_out = "downstream daily minimum DO")) %>%
  mutate(conf = fct_relevel(conf, "fon_vio", "lig_viz", "mou_fon", "loi_doi",
                             "mar_cur", "char_car", "coise_pot", "char_mou",
                             "coise_rieu", "coise_coiz")) %>%
  ggplot(aes(x = conf,    #fct_reorder(conf, value, .fun = function(.x) mean(.x == "in")),
             fill = value)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(type = "qual",
                    name = "",
                    labels = c("above upstream range",
                               "below upstream range",
                               "within upstream range")) +
  facet_wrap(~name) +
  theme_bw() +
  ylab("percentage of days") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5),
        axis.title.x = element_blank())
```

We see that for the all sites, less than half of the days with coincident 
upstream-downstream measurements have the downstream maximum constrained by
upstream measurements. This is similar for daily minima.

We can be slightly more quantitative by calculating the relative difference between 
upstream and downstream values, again focusing on daily maxima and minima. We further
focus only on days when the downstream maximum was GREATER than the upstream maxima,
and days when the downstream minima was LESS than the upstream minima.


```{r, echo=FALSE}
# Plot it
df_test %>%
  select(date, conf, per_ab, per_be) %>%
  pivot_longer(-c(conf, date)) %>%
  unnest() %>%
  drop_na() %>%
  filter(value > -0.80) %>%
  mutate(conf = fct_relevel(conf, "fon_vio", "lig_viz", "mou_fon", "loi_doi",
                            "mar_cur", "char_car", "coise_pot", "char_mou",
                            "coise_rieu", "coise_coiz")) %>%
  ggplot(aes(x = conf,
             y = value,
             fill = name)) +
  geom_boxplot() +
  geom_point(shape = 21, position = position_jitterdodge()) +
  scale_y_continuous(labels = scales::percent,
                     breaks = seq(-0.75, 0.5, 0.25)) +
  scale_fill_brewer(type = "qual",
                    name = "",
                    labels = c("% > upstream max.",
                               "% < upstream min.")) +
  theme_bw() +
  ylab("downstream-upstream % difference") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5),
        axis.title.x = element_blank())
```

It looks like most differences are in the 5–15% range, but there does appear to be
a decreasing trend in a "multiplicative effect" of daily maxima as 
specific conducivity differences decrease.


Do these differences happen in particular times of the year as opposed to others?
```{r, echo=FALSE}
# Plot it
df_test %>%
  select(date, conf, min_out, max_out) %>%
  pivot_longer(-c(conf,date)) %>%
  filter(year(date) == 2020,
         ((name == "min_out" & value == "below") |
            (name == "max_out" & value == "above"))) %>%
  group_by(conf, name) %>%
  arrange(date, value) %>%
  mutate(rn = row_number()) %>%
  ggplot(aes(x = date, y = rn,
             color = value)) +
  geom_step() +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  # scale_y_continuous(labels = scales::percent) +
  scale_color_brewer(type = "qual",
                    name = "",
                    labels = c("downstream max.> \n upstream range",
                               "downstream min.<  \n upstream range")) +
  facet_wrap(~conf) +
  theme_bw() +
  ylab("cumulative count of days in 2020") + xlab("")
```

Hard to tell. Below are the two confluences with the greatest SC differences,
and you can look at specific dates and times by zooming. It can be be a bit slow 
to respond because there is so much data, but it will get there after a few seconds.


```{r, echo=FALSE}
plot_ly(
  filter(df_DO, conf == "fon_vio"),               #indicate the dataframe you're using
  x = ~datetime,         #identify x variable (keep the tilda)
  y = ~DO,           #identify y variable (keep the tilda)
  text = ~site,
  color = ~site, #when hovering over a bubble will display this var.
  type = "scatter",   #makes the plot a scatterplot
  mode = "markers",   #indicates values are points (e.g. not lines)
  marker = list(      #chooses properties of each bubble
    opacity = 0.8)) %>%   #the smaller opacity the more see the bubble
  layout(title = 'Fontbonne-Violay', #names plot
         xaxis = list(showgrid = TRUE),       #makes x gridlines
         yaxis = list(showgrid = TRUE))

plot_ly(
  filter(df_DO, conf == "lig_viz"),               #indicate the dataframe you're using
  x = ~datetime,         #identify x variable (keep the tilda)
  y = ~DO,           #identify y variable (keep the tilda)
  text = ~site,
  color = ~site, #when hovering over a bubble will display this var.
  type = "scatter",   #makes the plot a scatterplot
  mode = "markers",   #indicates values are points (e.g. not lines)
  marker = list(      #chooses properties of each bubble
    opacity = 0.8)) %>%   #the smaller opacity the more see the bubble
  layout(title = 'Lignon-Vizézy', #names plot
         xaxis = list(showgrid = TRUE),       #makes x gridlines
         yaxis = list(showgrid = TRUE))

```

<!-- ```{r, echo=FALSE} -->
<!-- # match location with DO summary, summarize by week -->
<!-- do_loc <- df_DO %>% -->
<!--   mutate(hour = floor_date(datetime, "1 hour")) %>% -->
<!--   group_by(watershed, site, Longitude, Latitude, hour) %>% -->
<!--   summarize(DO = mean(DO, na.rm = TRUE)) %>% -->
<!--   ungroup() %>% -->
<!--   filter(hour > ymd("2020-04-24")) %>% -->
<!--   filter(hour < ymd("2020-04-29")) %>% -->
<!--   st_as_sf(coords = c("Longitude", "Latitude")) -->

<!-- # Animated map of GPP -->
<!-- anim <- ggplot() +  -->
<!--   geom_sf(data = filter(do_loc, watershed =="Loise"),#slice_sample(df_map, n = 5000), -->
<!--           aes(color = DO, -->
<!--               size = DO)) + -->
<!--   scale_color_viridis_c() + -->
<!--   transition_states(hour, -->
<!--                     transition_length = 1, -->
<!--                     state_length = 1) + -->
<!--   ggtitle('{closest_state}', -->
<!--           subtitle = 'Frame {frame} of {nframes}') -->
<!-- # anim -->
<!-- animate( -->
<!--   anim, -->
<!--   nframes = 300 -->
<!-- ) -->
<!-- ``` -->


<!-- ```{r, echo=FALSE} -->
<!-- # match location with DO summary, summarize by week -->
<!-- do_loc <- df_DO %>% -->
<!--   mutate(hour = floor_date(datetime, "1 hour")) %>% -->
<!--   group_by(watershed, site, Longitude, Latitude, hour) %>% -->
<!--   summarize(DO = mean(DO, na.rm = TRUE)) %>% -->
<!--   ungroup() %>% -->
<!--   filter(hour > ymd("2020-04-24")) %>% -->
<!--   filter(hour < ymd("2020-04-29")) %>% -->
<!--   st_as_sf(coords = c("Longitude", "Latitude")) -->

<!-- # Animated map of GPP -->
<!-- anim <- ggplot() +  -->
<!--   geom_sf(data = filter(do_loc, watershed =="Coise"),#slice_sample(df_map, n = 5000), -->
<!--           aes(color = DO, -->
<!--               size = DO)) + -->
<!--   scale_color_viridis_c() + -->
<!--   transition_states(hour, -->
<!--                     transition_length = 1, -->
<!--                     state_length = 1) + -->
<!--   ggtitle('{closest_state}', -->
<!--           subtitle = 'Frame {frame} of {nframes}') -->
<!-- # anim -->
<!-- animate( -->
<!--   anim, -->
<!--   nframes = 300 -->
<!-- ) -->
<!-- ``` -->